# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

#### Loading Data: activity data is unzipped and placed in ./activity/ directory
```{r ECHO=T}
fd <- './activity/'
fp <- paste(fd,'activity.csv', sep='')
activity <-read.csv(file=fp)
head(activity)

library(lubridate)
activity$date<- ymd(activity$date)
```

## What is mean total number of steps taken per day?
#### Summary by Date
```{r ECHO=T}
sumStepByDate <- data.frame(with(activity, tapply(steps, date, sum, na.rm=T, simplify = T)))
names(sumStepByDate) <- c('TotalSteps')
sumStepByDate$Date <- rownames(sumStepByDate)
library(lubridate)
sumStepByDate$Date <- ymd(sumStepByDate$Date)
str(sumStepByDate$Date)
```
#### Histogram of Total Number of Steps Taken Per Day
```{r ECHO=T}
hist(x=sumStepByDate$TotalSteps, main='Distribution of Total Number\n of Steps Taken Each Day')
```

#### Mean and Median Total Number of Steps Taken Per Day
```{r ECHO=T}
# Mean Value
mean(x=sumStepByDate$TotalSteps, na.rm = T)
# Median Value
median(x=sumStepByDate$TotalSteps, na.rm = T)
```


## What is the average daily activity pattern?
#### Summary by Interval
```{r ECHO=T}
AverageStepByInterval <- data.frame(with(activity, tapply(steps, interval, mean, na.rm=T, simplify = T)))
names(AverageStepByInterval) <- c('AverageSteps')
AverageStepByInterval$interval <- rownames(AverageStepByInterval)
str(AverageStepByInterval$interval)
```

#### Time Series Plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r ECHO=T}
with(AverageStepByInterval, plot(AverageSteps~interval, type="l"))
```

#### 5-minute interval contains the maximum number of steps
```{r ECHO=T}
ind<-with(AverageStepByInterval, which(AverageSteps == max(AverageSteps)))
AverageStepByInterval[ind,]
```

## Imputing missing values

#### Total Number of Missing Values
```{r ECHO=T}
cntNA = sum(is.na(activity$steps))
allCnt = nrow(activity)
rNA = cntNA / allCnt
sprintf("Number of Missing %d    Portion of Missing Observations %f", cntNA, rNA)
```

#### Filling missing Data using average of 5-minute interval
```{r ECHO=T}
activity_cleaned <- merge(activity, AverageStepByInterval, by.x="interval", by.y = "interval", all=T)
activity_cleaned$steps[is.na(activity_cleaned$steps)] <- activity_cleaned$AverageSteps[is.na(activity_cleaned$steps)]
activity_cleaned<-activity_cleaned[,c(1,2,3)]
head(activity_cleaned)
```

#### Summary by Date
```{r ECHO=T}
sumStepByDate <- data.frame(with(activity_cleaned, tapply(steps, date, sum, na.rm=T, simplify = T)))
names(sumStepByDate) <- c('TotalSteps')
sumStepByDate$Date <- rownames(sumStepByDate)
library(lubridate)
sumStepByDate$Date <- ymd(sumStepByDate$Date)
str(sumStepByDate$Date)
```

#### Histogram of Total Number of Steps Taken Per Day
```{r ECHO=T}
hist(x=sumStepByDate$TotalSteps, main='Distribution of Total Number\n of Steps Taken Each Day')
```

#### Mean and Median Total Number of Steps Taken Per Day
```{r ECHO=T}
# Mean Value
mean(x=sumStepByDate$TotalSteps, na.rm = F)
# Median Value
median(x=sumStepByDate$TotalSteps, na.rm = F)
```


## Are there differences in activity patterns between weekdays and weekends?
#### Assign new factor variable in the dataset activity_clean with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend
```{r ECHO=T}
isWeekday <- wday(activity_cleaned$date) > 1 & wday(activity_cleaned$date) < 7
activity_cleaned$daytype[isWeekday]<-'weekday'
activity_cleaned$daytype[!isWeekday]<-'weekend'
activity_cleaned$daytype <- as.factor(activity_cleaned$daytype)
```
####  time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r ECHO=T}
#### Summary by Interval
activity_avg_interval<-with(activity_cleaned, aggregate(steps, by=list(interval, daytype), FUN=mean))
names(activity_avg_interval) <- c('interval', 'daytype', 'avgsteps')
```
```{r ECHO=T}
library(lattice)
## Plot with 2 Panels
with (activity_avg_interval, xyplot(avgsteps~interval | daytype, layout = c(1,2), type="l",ylab="Number of steps"))
```